---
title: "Project Write-up"
output: html_document
---

First I load the necessary packages & download the datasets:
```{r,warnings=FALSE,message=FALSE,cache=TRUE}
library(dplyr)
library(caret)
library(randomForest)

trainURL <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-training.csv"
testURL <- "https://d396qusza40orc.cloudfront.net/predmachlearn/pml-testing.csv"

training <- read.csv(trainURL,na.strings=c("NA","","NaN"))
testing <- read.csv(testURL,na.strings=c("NA","","NaN"))
```

Second, I narrowed the dataset down to features that I thought would be most relevant to consider for model selection.  I first removed columns that would probably not be useful when making out-of-sample predictions such as row index, user_names, timestamps and derived measures.  Then I removed columns whose proportion of missing data was > 97%.  Lastly, I removed columns that were highly correlated above a threshold of 0.75.  This left me with a more reasonable number of predictors (31) rather than the unwieldy 160 columns in the original datasets.

```{r,cache=TRUE}
# Remove non-numerics and timestamp data
training <- training[,8:160]

# Explore number of missing for each predictor
nas <- sapply(as.data.frame(sapply(training,is.na)),sum)

# Remove predictors that are mostly missing
training <- training[,which(nas < 19216)]

# Find and remove highly correlated
nums <- sapply(training,is.numeric)
filtered.nums <- training[complete.cases(training),nums]
corrs <- cor(filtered.nums)
highlycorr <- findCorrelation(corrs,cutoff=0.75)

corr.names <- names(filtered.nums)[highlycorr]

filtered <- training[,-which(names(training) %in% corr.names)]
```

I trained a random forest model with 3-fold cross-validation.

```{r,cache=TRUE}
# Train RF model with 3-fold cross-validation
set.seed(1981)
model <- train(classe~.,data=filtered,method="rf",trControl=trainControl(method="cv",number=3))

model
model$finalModel
```

The model was highly accurate with m=16 yielding an accuracy of >99% and an OOB estimate of error of 0.55%.

```{r,echo=FALSE}
test_preds <- predict(model,testing)
for (i in 1:length(test_preds)){
    filename <- paste0("problem_id_",i,".txt")
    write.table(test_preds[i],file=filename,quote=FALSE,row.names=FALSE,col.names=FALSE)
}
```

